### 2024-11-27 doc.tex
Start of report

\documentclass[11pt]{article}
\usepackage[T1]{fontenc}
\usepackage{libertine}
\renewcommand{\sfdefault}{fvs} % Use as sans-serif
\renewcommand{\ttdefault}{fvm} % Use as monospace
%\usepackage{helvet} % Helvetica font for sans-serif
%\usepackage{courier} % Courier font for monospace
\usepackage{hyperref}
\usepackage[backend=biber,style=authoryear,autocite=inline]{biblatex}
\usepackage{titlesec}

\titleformat{\section}{\sffamily\Large}{\thesection}{1em}{}
\titleformat{\subsection}{\sffamily\large}{\thesubsection}{1em}{}
\titleformat{\subsubsection}{\sffamily\normalsize}{\thesubsubsection}{1em}{}

\addbibresource{ref.bib}

\begin{document}

\section*{AI energy consumption}

\begin{itemize}
    \item \cite{castano2023exploring}
    \item \cite{devries2023growing}
    \item \cite{heguerte2023estimate}
    \item \cite{gowda2024watt}
    \item \cite{luccioni2023counting}
    \item \cite{patterson2021carbon}
    \item \cite{rodriguez2024evaluating}
    \item \cite{tomlinson2024carbon}
    \item \cite{tripp2024measuring}
\end{itemize}

xxxx

\section*{Energy mix}
\begin{itemize}
    \item \cite{krebs2018umweltbilanz}
\end{itemize}

\printbibliography
\end{document}